---
editor_options:
  chunk_output_type: console
---

# NOTES

There's some sparse documentation available for download (PDF)
[here](https://www.bfs.admin.ch/bfs/de/home/dienstleistungen/forschung/api/api-gemeinde.assetdetail.15224054.html).

## API response's raw column names

An example snippet to determine them for the `snapshots` endpoint:

```{r, purl = FALSE}
httr::RETRY(verb = "GET",
            url = api_base_url,
            path = paste0(api_common_path, "snapshots"),
            query = list(startPeriod = NULL,
                         endPeriod = NULL,
                         useBfsCode = tolower(TRUE))) %>%
  httr::content(as = "text") %>%
  stringr::str_split_1(pattern = "\n") %>%
  dplyr::first() %>%
  stringr::str_split_1(pattern = ",") %>%
  pal::cat_lines()
```

## API URL parameter details

### `historicized_code`

| column name if `historicized_code = TRUE` | column name if `historicized_code = FALSE` |
|-------------------------------------------|--------------------------------------------|
| CODE_HIST_1\_Text_en                      | CODE_OFS_1\_Text_en                        |
| CODE_HIST_1\_Text_fr                      | CODE_OFS_1\_Text_fr                        |
| CODE_HIST_1\_Text_de                      | CODE_OFS_1\_Text_de                        |
| CODE_HIST_1\_Text_it                      | CODE_OFS_1\_Text_it                        |
| CODE_HIST_1\_Text                         | CODE_OFS_1\_Text                           |

-   The `CODE_OFS*` corresponds to the [official Swiss community identification
    number](https://en.wikipedia.org/wiki/Community_Identification_Number#Switzerland) (also called "GEOSTAT"/"BFS" number, see the [corresponding German
    Wikipedia article](https://de.wikipedia.org/wiki/Gemeindenummer)).
-   The `CODE_HIST*` corresponds to the "historical number". -\> The URL parameter's name `useBfsCode` seems misleading (inversed)!

### `format`

-   Setting `format=csv` or none results in the normal CSV response.

-   Setting `format` to anything else than `csv` results in an **XLS response**!

    -   For the `snapshots`, `correspondances` and `mutations` endpoints, the XLS contains the normal data.
    -   For the `levels` endpoint, the XLS contains the **column name spec** of the normal API response's `HR_*` columns. The `levels` endpoint is also the only
        one offering the `labelLanguages` URL param (see below).

### `labelLanguages`

Only present in the `levels` API endpoint. Can be set to a comma-separated list of language abbreviations. Supported are `de`, `en`, `fr` and `it`.

### `oracleCompatible`

Not a clue what that is supposed to do. The following has been tried and *didn't* change the response:

-   `oracleCompatible`
-   `oracleCompatible=true`
-   `oracleCompatible=false`

# INTERNAL

## Avoid `R CMD check` notes about undefined global objects used in magrittr pipes

cf. <https://github.com/tidyverse/magrittr/issues/29#issuecomment-74313262>

```{r}
utils::globalVariables(names = ".")
```

## Package load/unload

```{r}
.onLoad <- function(libname, pkgname) {
  pkgpins::clear_cache(board = pkgpins::board(pkg = pkgname),
                       max_age = getOption(paste0(pkgname, ".max_cache_lifespan"),
                                           default = "30 days"))
}
```

## Constants

```{r}
this_pkg <- utils::packageName()

# URLs/paths common to all API endpoints
api_base_url <- "https://sms.bfs.admin.ch/"
api_common_path <- "WcfBFSSpecificService.svc/AnonymousRest/communes/"

# Unicode copy/paste symbol
cp_symbol <- "\u2398"
```

## Functions

### `as_api_date`

Convert ISO-8601 formatted date string to `DD-MM-YYYY` format

```{r}
as_api_date <- function(date) {
  
  lubridate::ymd(date) %>%
    checkmate::assert_date(any.missing = FALSE,
                           null.ok = TRUE,
                           .var.name = "start/end_date") %>%
    pal::when(length(.) == 0L ~ NULL,
              ~ format(., "%d-%m-%Y"))
}
```

### `api_params`

```{r}
#' Get allowed API URL parameters
#' 
#' Parses the API's [Web Services Description Language (WSDL)](https://en.wikipedia.org/wiki/Web_Services_Description_Language) file to determine
#' possible URL parameters.
#'
#' @param type The API endpoint.
#'
#' @return A [tibble][tibble::tbl_df].
#' @keywords internal
#' @export
#'
#' @examples
#' # get all URL params of all 3 API endpoints:
#' endpoints <- c("snapshots",
#'                "congruences",
#'                "mutations",
#'                "classifications")
#'
#' endpoints |>
#'   purrr::map(swissmuni::api_params) |>
#'   rlang::set_names(nm = endpoints)
api_params <- function(type = c("snapshots",
                                "congruences",
                                "mutations",
                                "classifications")) {
  
  pal::assert_pkg("xml2")
  
  rlang::arg_match(type) %>%
    switch("snapshots" = "GetSnapshot",
           "congruences" = "GetCorrespondances",
           "mutations" = "GetMutations",
           "classifications" = "GetGeographicLevel") %>%
    paste0("//xs:element[@name='", ., "']") %>%
    xml2::xml_find_all(x = xml2::read_xml("https://sms.bfs.admin.ch/WcfBFSSpecificService.svc?singleWsdl")) %>%
    checkmate::assert_list(len = 1L,
                           types = "xml_node",
                           any.missing = FALSE) %>%
    xml2::xml_find_all(xpath = ".//xs:element") %>%
    xml2::xml_attrs() %>%
    purrr::map(as.list) %>%
    purrr::map(tibble::as_tibble) %>%
    purrr::list_rbind()
}
```

### `parse_result`

NOTES:

-   The API response doesn't have the proper content type set (would be `text/csv`).

-   The API returns different numbers of columns depending on the provided URL params (like the `start/end_date`). Therefore we generate the column types
    telling `readr::read_csv()` how to parse the results dynamically.

```{r}
parse_result <- function(response,
                         col_types) {
  
  response %<>% httr::content(as = "text")
  
  response %>% readr::read_csv(col_types = pal::cols_regex("(Name|_Title|_Text|^ABBREV)" = readr::col_character(),
                                                           "^(MutationDate|ValidFrom|ValidTo)$" = readr::col_date(format = "%d.%m.%Y"),
                                                           .col_names = pal::dsv_colnames(response),
                                                           .default = readr::col_integer()))
}
```

# EXPORTED

## Municipality data

### `snapshots`

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/snapshots?useBfsCode=true&startPeriod=01-01-2018&endPeriod=01-01-2018>

Possible URL parameters:

-   `format`
-   `escapeChars`
-   `startPeriod`
-   `endPeriod`
-   `useBfsCode`

```{r}
#' Get municipality snapshots
#'
#' Returns a [tibble][tibble::tbl_df] with snapshots of Swiss municipality data. A snapshot corresponds to the list of municipalities that are present at least
#' partially during the specified period (or at the specified day when `end_date = start_date`).
#'
#' @includeRmd snippets/fso_vs_historicized_code.Rmd
#'
#' @param start_date `r pkgsnip::param_label("start_date")`
#' @param end_date `r pkgsnip::param_label("end_date")`
#' @param historicized_code By default, the [_FSO commune number_](https://de.wikipedia.org/wiki/Gemeindenummer) is returned. Set to `TRUE` in order to get the
#'   _historicization number_ instead.
#' @param use_cache `r pkgsnip::param_label("use_cache")`
#' @param cache_lifespan `r pkgsnip::param_label("cache_lifespan")` Defaults to 1 day (24 hours).
#'
#' @return `r pkgsnip::return_label("data")`
#' @export
#'
#' @examples
#' swissmuni::snapshots(historicized_code = TRUE)
snapshots <- function(start_date = lubridate::today(),
                      end_date = start_date,
                      historicized_code = FALSE,
                      use_cache = TRUE,
                      cache_lifespan = "1 day") {
  
  pkgpins::with_cache(expr = {
    
    httr::RETRY(verb = "GET",
                url = api_base_url,
                path = paste0(api_common_path, "snapshots"),
                query = list(startPeriod = as_api_date(start_date),
                             endPeriod = as_api_date(end_date),
                             useBfsCode = tolower(checkmate::assert_flag(historicized_code))),
                times = 5L) %>%
      parse_result()
  },
  pkg = this_pkg,
  from_fn = "snapshots",
  start_date,
  end_date,
  historicized_code,
  use_cache = use_cache,
  cache_lifespan = cache_lifespan)
}
```

### `congruences`

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/correspondances?includeUnmodified=true&startPeriod=01-01-2017&endPeriod=01-01-2018>

Possible URL parameters:

-   `format`
-   `escapeChars`
-   `startPeriod`
-   `endPeriod`
-   `includeUnmodified`
-   `includeTerritoryExchange`

```{r}
#' Get municipality congruences
#'
#' Returns a [tibble][tibble::tbl_df] with congruence data of Swiss municipalities. The data indicates for each existing municipality at the `start_date` to
#' which municipality/municipalities it corresponds at the `end_date`.
#'
#' @inheritParams snapshots
#' @param incl_unmodified Include municipalities where no changes occurred in the defined period.
#' @param incl_territory_exchange Include information (additional rows) about territory exchanges within and between municipalities.
#'
#' @inherit snapshots return
#' @export
#'
#' @examples
#' swissmuni::congruences(start_date = lubridate::today(),
#'                        end_date = lubridate::today(),
#'                        cache_lifespan = "6 hours")
congruences <- function(start_date = NULL,
                        end_date = NULL,
                        incl_unmodified = TRUE,
                        incl_territory_exchange = FALSE,
                        use_cache = TRUE,
                        cache_lifespan = "1 day") {
  
  pkgpins::with_cache(expr = {
    
    httr::RETRY(verb = "GET",
                url = api_base_url,
                path = paste0(api_common_path, "correspondances"),
                query = list(startPeriod = as_api_date(start_date),
                             endPeriod = as_api_date(end_date),
                             includeUnmodified = tolower(checkmate::assert_flag(incl_unmodified)),
                             includeTerritoryExchange = tolower(checkmate::assert_flag(incl_territory_exchange))),
                times = 5L) %>%
      parse_result()
  },
  pkg = this_pkg,
  from_fn = "congruences",
  start_date,
  end_date,
  incl_unmodified,
  incl_territory_exchange,
  use_cache = use_cache,
  cache_lifespan = cache_lifespan)
}
```

### `mutations`

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/mutations?startPeriod=01-01-2017&endPeriod=01-01-2018&includeTerritoryExchange=false>

Possible URL parameters:

-   `format`
-   `escapeChars`
-   `startPeriod`
-   `endPeriod`
-   `includeTerritoryExchange`

```{r}
#' Get municipality mutations
#'
#' Returns a [tibble][tibble::tbl_df] with mutation data of Swiss municipalities. The data describes all changes related to the municipalities made between
#' `start_date` and `end_date`.
#'
#' @inheritParams snapshots
#' @inheritParams congruences
#'
#' @inherit snapshots return
#' @export
#'
#' @examples
#' swissmuni::mutations(start_date = "2020-01-01",
#'                      end_date = "2020-06-30",
#'                      incl_territory_exchange = TRUE)
mutations <- function(start_date = NULL,
                      end_date = NULL,
                      incl_territory_exchange = FALSE,
                      use_cache = TRUE,
                      cache_lifespan = "1 day") {
  
  pkgpins::with_cache(expr = {
    
    httr::RETRY(verb = "GET",
                url = api_base_url,
                path = paste0(api_common_path, "mutations"),
                query = list(startPeriod = as_api_date(start_date),
                             endPeriod = as_api_date(end_date),
                             includeTerritoryExchange = tolower(checkmate::assert_flag(incl_territory_exchange))),
                times = 5L) %>%
      parse_result()
  },
  pkg = this_pkg,
  from_fn = "mutations",
  start_date,
  end_date,
  incl_territory_exchange,
  use_cache = use_cache,
  cache_lifespan = cache_lifespan)
}
```

## Municipality metadata

### `classifications`

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/levels?startPeriod=01-01-2017&endPeriod=01-01-2018>

Possible URL parameters:

-   `format`
-   `escapeChars`
-   `startPeriod`
-   `endPeriod`
-   `useBfsCode`
-   `labelLanguages`
-   `oracleCompatible`

```{r}
#' Get spatial classifications of municipalities
#'
#' Returns a [tibble][tibble::tbl_df] with a wide range of spatial classifications of Swiss municipalities, i.a. by language region, size of the municipality or
#' degree of agglomeration/urbanization.
#'
#' There are some classifications which have multiple column IDs (e.g. `HR_ARBREG2000_L2` and `HR_TYPMSR2000_L2` both contain the _spatial mobility regions_).
#' If `name_type` is set to a language code, column duplicates are given the name suffix "`r paste0(' (', cp_symbol, ')')`".
#'
#' @includeRmd snippets/fso_vs_historicized_code.Rmd
#'
#' @inheritParams snapshots
#' @param name_type The column name type to be returned. Defaults to the column IDs returned by the web service. If set to a language code, the columns
#'   starting with ID `HR_` will be named by its full label in the respective language instead.
#'
#' @inherit snapshots return
#' @export
#'
#' @examples
#' swissmuni::classifications(start_date = lubridate::today(),
#'                            end_date = lubridate::today(),
#'                            name_type = "de",
#'                            use_cache = FALSE)
classifications <- function(start_date = NULL,
                            end_date = NULL,
                            historicized_code = FALSE,
                            name_type = c("ID", "en", "de", "fr", "it"),
                            use_cache = TRUE,
                            cache_lifespan = "1 day") {
  
  result <- pkgpins::with_cache(
    expr = {
      
      httr::RETRY(verb = "GET",
                  url = api_base_url,
                  path = paste0(api_common_path, "levels"),
                  query = list(startPeriod = as_api_date(start_date),
                               endPeriod = as_api_date(end_date),
                               useBfsCode = tolower(checkmate::assert_flag(historicized_code))),
                  times = 5L) %>%
        parse_result()
    },
    pkg = this_pkg,
    from_fn = "classifications",
    start_date,
    end_date,
    historicized_code,
    use_cache = use_cache,
    cache_lifespan = cache_lifespan
  )
  
  name_type <- rlang::arg_match(name_type)
  
  if (name_type != "ID") {
    
    hr_ix <- which(stringr::str_detect(colnames(result), "^HR_"))
    hr_ids <- colnames(result) %>% stringr::str_subset("^HR_")
    hr_names <- col_spec_classifications[[paste0("name_", name_type)]][match(hr_ids, col_spec_classifications$column_id)]
    
    # ensure column names are unique (there are different column IDs for the same values)
    # -> add the unicode symbol for copy/paste, multiple times if necessary
    while (anyDuplicated(hr_names)) {
      hr_names[which(duplicated(hr_names))] <- paste0(hr_names[which(duplicated(hr_names))], " ", cp_symbol)
    }
    
    colnames(result)[hr_ix] <- hr_names
  }
  
  result
}
```
