---
editor_options:
  chunk_output_type: console
---

# INTERNAL

## Avoid `R CMD check` notes about undefined global objects used in magrittr pipes

cf. <https://github.com/tidyverse/magrittr/issues/29#issuecomment-74313262>

```{r}
utils::globalVariables(names = c("."))
```

## Package load/unload

```{r}
.onLoad <- function(libname, pkgname) {
  pkgpins::clear(pkg = pkgname)
}

.onUnload <- function (libpath) {
  pkgpins::deregister(pkg = pkg)
}
```

## Constants

```{r}
pkg <- utils::packageName()

# URLs/paths common to all API endpoints
api_base_url <- "https://sms.bfs.admin.ch/"
api_common_path <- "WcfBFSSpecificService.svc/AnonymousRest/communes/"

# CSV column names which should be parsed as dates in the format `DD.MM.YYYY`
date_colnames <- c("MutationDate",
                   "ValidFrom",
                   "ValidTo")

# WSDL XML
api_wsdl <- xml2::read_xml("https://sms.bfs.admin.ch/WcfBFSSpecificService.svc?singleWsdl")
```

## Convert ISO-8601 formatted date string to `DD-MM-YYYY` format

```{r}
as_api_date <- function(date) {
  
  lubridate::ymd(date) %>%
    checkmate::assert_date(any.missing = FALSE,
                           null.ok = TRUE,
                           .var.name = "start/end_date") %>%
    purrr::when(length(.) == 0 ~ NULL,
                ~ format(., "%d-%m-%Y"))
}
```

## Get allowed API URL parameters

This function parses the API's [Web Services Description Language (WSDL)](https://en.wikipedia.org/wiki/Web_Services_Description_Language) file to determine
possible URL parameters.

Use the following code to get a tibble per API endpoint containing param details:

```r
c("snapshots",
  "congruences",
  "mutations",
  "classifications") %>%
  {
    rlang::set_names(x = purrr::map(.x = .,
                                    .f = swissmuni:::api_params),
                     nm = .)
  }
```

```{r}
#' Get allowed API URL parameters
#'
#' @param type The API endpoint.
#'
#' @return A [tibble][tibble::tbl_df].
#' @keywords internal
#'
#' @examples
#' # to get all params for all 3 API endpoints:
#' library(magrittr)
#'
#' endpoints <- c("snapshots",
#'                "congruences",
#'                "mutations",
#'                "classifications")
#'
#' \dontrun{
#' endpoints %>%
#'   purrr::map(api_params) %>%
#'   rlang::set_names(nm = endpoints)}
api_params <- function(type = c("snapshots",
                                "congruences",
                                "mutations",
                                "classifications")) {
  
  rlang::arg_match(type) %>%
    switch("snapshots" = "GetSnapshot",
           "congruences" = "GetCorrespondances",
           "mutations" = "GetMutations",
           "classifications" = "GetGeographicLevel") %>%
    paste0("//xs:element[@name='", ., "']") %>%
    xml2::xml_find_all(x = api_wsdl) %>%
    checkmate::assert_list(len = 1,
                           types = "xml_node",
                           any.missing = FALSE) %>%
    xml2::xml_find_all(xpath = ".//xs:element") %>%
    xml2::xml_attrs() %>%
    purrr::map(as.list) %>%
    purrr::map_dfr(tibble::as_tibble)
}
```

## Parse API result

**Notes:**

- The API doesn't have the proper content type set (would be `text/csv`).

```{r}
parse_result <- function(response,
                         col_types) {
  
  response %>%
    httr::content(as = "text") %>%
    readr::read_csv(col_types = col_types) %>%
    dplyr::mutate(dplyr::across(dplyr::any_of(date_colnames),
                                lubridate::dmy))
}
```

# Get municipality snapshots

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/snapshots?useBfsCode=true&startPeriod=01-01-2018&endPeriod=01-01-2018>

Possible URL parameters:

- `format`
- `escapeChars`
- `startPeriod`
- `endPeriod`
- `useBfsCode`

---

Differences regarding URL parameters:

column name if `use_fso_code = TRUE` | column name if `use_fso_code = FALSE`
------------------------------------ | -------------------------------------
CODE_HIST_1_Text_en                  | CODE_OFS_1_Text_en
CODE_HIST_1_Text_fr                  | CODE_OFS_1_Text_fr
CODE_HIST_1_Text_de                  | CODE_OFS_1_Text_de
CODE_HIST_1_Text_it                  | CODE_OFS_1_Text_it
CODE_HIST_1_Text                     | CODE_OFS_1_Text

The `CODE_OFS*` corresponds to the ["GEOSTAT"/"BFS" number](https://de.wikipedia.org/wiki/Gemeindenummer) and the `CODE_HIST*` to the "historical number".
-> The URL parameter's name `useBfsCode` seems misleading (inversed)!

---

Determine the column names:

```r
httr::GET(url = api_base_url,
                path = paste0(api_common_path, "snapshots"),
                query = list(startPeriod = as_api_date(NULL),
                             endPeriod = as_api_date(NULL),
                             useBfsCode = tolower(T))) %>%
  httr::content(as = "text") %>%
  stringr::str_split(pattern = "\n") %>%
  dplyr::first() %>%
  dplyr::first() %>%
  stringr::str_split(pattern = ",") %>%
  pal::cat_lines()
```

```{r}
#' Get municipality snapshots
#'
#' This function returns a [tibble][tibble::tbl_df] with snapshots of Swiss municipality data. A snapshot is the state of the municipalities at a specified
#' point in time. Therefore you probably want to set `start_date = end_date`.
#'
#' @includeRmd snippets/fso_vs_historicized_code.Rmd
#'
#' @param start_date The begin of the period the data covers. A [date](base::Date) or a character scalar in the format `"YYYY-MM-DD"`.
#' @param end_date The end of the period the data covers. A [date](base::Date) or a character scalar in the format `"YYYY-MM-DD"`.
#' @param historicized_code If `TRUE`, the `CODE_HIST` of the municipalities is returned instead of the `CODE_OFS`?
#' @param use_cache `r pkgsnippets::param_label("use_cache")`
#' @param cache_lifespan `r pkgsnippets::param_label("cache_lifespan")` Defaults to 1 day (24 hours).
#'
#' @return `r pkgsnippets::return_label("data")`
#' @export
#'
#' @examples
#' snapshots(start_date = "2020-01-01",
#'           end_date = "2020-01-01",
#'           use_cache = FALSE)
snapshots <- function(start_date = lubridate::today() - 1,
                      end_date = start_date,
                      historicized_code = FALSE,
                      use_cache = TRUE,
                      cache_lifespan = "1 day") {
  
  if (checkmate::assert_flag(use_cache)) {
    
    pin_name <- pkgpins::call_to_name()
    result <- pkgpins::get_obj(id = pin_name,
                               max_age = cache_lifespan,
                               pkg = pkg)
    fetch <- is.null(result)
    
  } else {
    fetch <- TRUE
  }
  
  if (fetch) {
    
    result <-
      httr::GET(url = api_base_url,
                path = paste0(api_common_path, "snapshots"),
                query = list(startPeriod = as_api_date(start_date),
                             endPeriod = as_api_date(end_date),
                             useBfsCode = tolower(historicized_code))) %>%
      parse_result(col_types = pal::as_string("iccii", rep("c",
                                                           times = 29)))
  }
  
  if (use_cache) {
    
    pkgpins::cache_obj(x = result,
                       id = pin_name,
                       pkg = pkg)
  }
  
  result
}
```

# Get municipality congruences

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/correspondances?includeUnmodified=true&startPeriod=01-01-2017&endPeriod=01-01-2018>

Possible URL parameters:

- `format`
- `escapeChars`
- `startPeriod`
- `endPeriod`
- `includeUnmodified`
- `includeTerritoryExchange`

```{r}
#' Get municipality congruences
#'
#' This function returns a [tibble][tibble::tbl_df] with congruence data of Swiss municipalities.
#'
#' @inheritParams snapshots
#' @param incl_unmodified Include municipalities where no changes occurred in the defined period.
#' @param incl_territory_exchange Include information about territory exchanges between municipalities.
#'
#' @inherit snapshots return
#' @export
#'
#' @examples
#' congruences(start_date = "2010-01-01",
#'             end_date = Sys.Date(),
#'             cache_lifespan = "6 hours")
congruences <- function(start_date = NULL,
                        end_date = NULL,
                        incl_unmodified = TRUE,
                        incl_territory_exchange = FALSE,
                        use_cache = TRUE,
                        cache_lifespan = "1 day") {
  
  if (checkmate::assert_flag(use_cache)) {
    
    pin_name <- pkgpins::call_to_name()
    result <- pkgpins::get_obj(id = pin_name,
                               max_age = cache_lifespan,
                               pkg = pkg)
    fetch <- is.null(result)
    
  } else {
    fetch <- TRUE
  }
  
  if (fetch) {
    
    result <-
      httr::GET(url = api_base_url,
                path = paste0(api_common_path, "correspondances"),
                query = list(startPeriod = as_api_date(start_date),
                             endPeriod = as_api_date(end_date),
                             includeUnmodified = tolower(incl_unmodified),
                             includeTerritoryExchange = tolower(incl_territory_exchange))) %>%
      parse_result(col_types = "iiciciiicici")
  }
  
  if (use_cache) {
    
    pkgpins::cache_obj(x = result,
                       id = pin_name,
                       pkg = pkg)
  }
  
  result
}
```

# Get municipality mutations

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/mutations?startPeriod=01-01-2017&endPeriod=01-01-2018&includeTerritoryExchange=false>

Possible URL parameters:

- `format`
- `escapeChars`
- `startPeriod`
- `endPeriod`
- `includeTerritoryExchange`

```{r}
#' Get municipality mutations
#'
#' This function returns a [tibble][tibble::tbl_df] with mutation data of Swiss municipalities.
#'
#' @inheritParams snapshots
#' @inheritParams congruences
#'
#' @inherit snapshots return
#' @export
#'
#' @examples
#' mutations(start_date = "2020-01-01",
#'           end_date = "2020-06-01")
mutations <- function(start_date = NULL,
                      end_date = NULL,
                      incl_territory_exchange = FALSE,
                      use_cache = TRUE,
                      cache_lifespan = "1 day") {
  
  if (checkmate::assert_flag(use_cache)) {
    
    pin_name <- pkgpins::call_to_name()
    result <- pkgpins::get_obj(id = pin_name,
                               max_age = cache_lifespan,
                               pkg = pkg)
    fetch <- is.null(result)
    
  } else {
    fetch <- TRUE
  }
  
  if (fetch) {
    
    result <-
      httr::GET(url = api_base_url,
                path = paste0(api_common_path, "mutations"),
                query = list(startPeriod = as_api_date(start_date),
                             endPeriod = as_api_date(end_date),
                             includeTerritoryExchange = tolower(incl_territory_exchange))) %>%
      parse_result(col_types = "iciiciciiicici")
  }
  
  if (use_cache) {
    
    pkgpins::cache_obj(x = result,
                       id = pin_name,
                       pkg = pkg)
  }
  
  result
}
```

# Get geographical classifications of municipalities

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/levels?startPeriod=01-01-2017&endPeriod=01-01-2018>

Possible URL parameters:

- `format`
- `escapeChars`
- `startPeriod`
- `endPeriod`
- `useBfsCode`
- `labelLanguages`[^no-clue]
- `oracleCompatible` [^no-clue]

[^no-clue]: Not a clue what that is supposed to do...

```{r}
#' Get geographical classifications of municipalities
#'
#' This function returns a [tibble][tibble::tbl_df] with a wide range of geographical classifications of Swiss municipalities.
#'
#' @includeRmd snippets/fso_vs_historicized_code.Rmd
#'
#' @inheritParams snapshots
#'
#' @inherit snapshots return
#' @export
#'
#' @examples
#' classifications(historicized_code = TRUE,
#'                 cache_lifespan = "1 week")
classifications <- function(start_date = NULL,
                            end_date = NULL,
                            historicized_code = FALSE,
                            use_cache = TRUE,
                            cache_lifespan = "1 day") {
  
  if (checkmate::assert_flag(use_cache)) {
    
    pin_name <- pkgpins::call_to_name()
    result <- pkgpins::get_obj(id = pin_name,
                               max_age = cache_lifespan,
                               pkg = pkg)
    fetch <- is.null(result)
    
  } else {
    fetch <- TRUE
  }
  
  if (fetch) {
    
    result <-
      httr::GET(url = api_base_url,
                path = paste0(api_common_path, "levels"),
                query = list(startPeriod = as_api_date(start_date),
                             endPeriod = as_api_date(end_date),
                             useBfsCode = tolower(historicized_code))) %>%
      parse_result(col_types = pal::as_string("iciii", rep("i",
                                                           times = 49), "ci"))
  }
  
  if (use_cache) {
    
    pkgpins::cache_obj(x = result,
                       id = pin_name,
                       pkg = pkg)
  }
  
  result
}
```
