---
editor_options:
  chunk_output_type: console
---

# NOTES

## API response's raw column names

An example snippet to determine them for the `snapshots` endpoint:

```{r, purl = FALSE}
httr::RETRY(verb = "GET",
            url = api_base_url,
            path = paste0(api_common_path, "snapshots"),
            query = list(startPeriod = NULL,
                         endPeriod = NULL,
                         useBfsCode = tolower(T))) %>%
  httr::content(as = "text") %>%
  stringr::str_split(pattern = "\n") %>%
  dplyr::first() %>%
  dplyr::first() %>%
  stringr::str_split(pattern = ",") %>%
  pal::cat_lines()
```

### Generate API response's full column names

- Unfortunately, this is only possible for the `HR_*` columns of the `levels` API endpoint.

- The returned names seem to be right (there are columns with completely different IDs like `HR_ARBREG2000_L2` and `HR_TYPMSR2000_L2` but identical values),
  except some year numbers. They are corrected by the function below.

- The `MS` in the full column names returned by `gen_col_spec_classifications()` stands for _mobilité spatiale_, cf.
  <https://www.bfs.admin.ch/bfs/de/home/statistiken/raum-umwelt/nomenklaturen/msreg.assetdetail.415729.html>

- The code chunk below writes the return value of `gen_col_spec_classifications()` to the package file `R/sysdata.rda`.

```{r, purl = FALSE}
gen_col_spec_classifications <- function() {
  
  # download raw XLS file
  tmp_file <- fs::file_temp(pattern = "col_spec_classifications")
  on.exit(unlink(tmp_file))
  
  httr::RETRY(verb = "GET",
              url = "https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/levels?format=x",
              httr::write_disk(path = tmp_file,
                               overwrite = TRUE),
              times = 5L)
  
  # tidy up data
  readxl::read_excel(path = tmp_file) %>%
    dplyr::rename_with(.fn =
                         ~ tolower(.x) %>%
                         stringr::str_remove_all(pattern = "[\\[\\]]") %>%
                         stringr::str_replace_all(pattern = "\\s+",
                                                  replacement = "_")) %>%
    ## complete duplicated names with missing year
    dplyr::rowwise() %>%
    dplyr::mutate(dplyr::across(dplyr::starts_with("name_"),
                                ~ if (stringr::str_detect(column_id, "^HR_(AGGLGK20(00|12)_L1|DEGURB|MSREG|SPRGEB|STALAN)")) {
                                  paste(.x, stringr::str_extract(column_id, "\\d{4}"))
                                } else .x)) %>%
    ## fix remaining duplicated names
    dplyr::mutate(dplyr::across(dplyr::starts_with("name_"),
                                ~ if (stringr::str_detect(column_id, "^HR_GDETYP(1980|1990|2000)_L[12]$")) {
                                  stringr::str_replace(.x, "1980-2000", stringr::str_extract(column_id, "\\d{4}"))
                                } else .x)) %>%
    dplyr::ungroup()
}

col_spec_classifications <- gen_col_spec_classifications()

usethis::use_data(col_spec_classifications,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "gzip", # difference to "xz" is only ~ +10%
                  version = 3L)
```

## API URL parameter details

### `historicized_code`:

column name if `historicized_code = TRUE` | column name if `historicized_code = FALSE`
----------------------------------------- | -------------------------------------
CODE_HIST_1_Text_en                       | CODE_OFS_1_Text_en
CODE_HIST_1_Text_fr                       | CODE_OFS_1_Text_fr
CODE_HIST_1_Text_de                       | CODE_OFS_1_Text_de
CODE_HIST_1_Text_it                       | CODE_OFS_1_Text_it
CODE_HIST_1_Text                          | CODE_OFS_1_Text

The `CODE_OFS*` corresponds to the ["GEOSTAT"/"BFS" number](https://de.wikipedia.org/wiki/Gemeindenummer) and the `CODE_HIST*` to the "historical number".
-> The URL parameter's name `useBfsCode` seems misleading (inversed)!

### `format`

- Setting `format=csv` or none results in the normal CSV response.

- Setting `format` to anything else than `csv` results in a **XLS response**!

  - For the `snapshots`, `correspondances` and `mutations` endpoints, the XLS contains the normal data.
  - For the `levels` endpoint, the XLS contains the **column name spec** of the normal API response's `HR_*` columns. The `levels` endpoint is also the only
    one offering the `labelLanguages` URL param (see below).

### `labelLanguages`

Only present in the `levels` API endpoint. Can be set to a comma-separated list of language abbreviations. Supported are `de`, `en`, `fr` and `it`.

### `oracleCompatible`

Not a clue what that is supposed to do. The following has been tried and _didn't_ change the response:

- `oracleCompatible`
- `oracleCompatible=true`
- `oracleCompatible=false`

# INTERNAL

## Avoid `R CMD check` notes about undefined global objects used in magrittr pipes

cf. <https://github.com/tidyverse/magrittr/issues/29#issuecomment-74313262>

```{r}
utils::globalVariables(names = ".")
```

## Package load/unload

```{r}
.onLoad <- function(libname, pkgname) {
  pkgpins::clear(pkg = pkgname,
                 max_age = getOption("swissmuni.max_cache_lifespan",
                                     default = "30 days"))
}

.onUnload <- function(libpath) {
  pkgpins::deregister(pkg = pkg)
}
```

## Constants

```{r}
pkg <- utils::packageName()

# URLs/paths common to all API endpoints
api_base_url <- "https://sms.bfs.admin.ch/"
api_common_path <- "WcfBFSSpecificService.svc/AnonymousRest/communes/"

# Unicode copy/paste symbol
cp_symbol <- "\u2398"
```

## Convert ISO-8601 formatted date string to `DD-MM-YYYY` format

```{r}
as_api_date <- function(date) {
  
  lubridate::ymd(date) %>%
    checkmate::assert_date(any.missing = FALSE,
                           null.ok = TRUE,
                           .var.name = "start/end_date") %>%
    purrr::when(length(.) == 0L ~ NULL,
                ~ format(., "%d-%m-%Y"))
}
```

## Get allowed API URL parameters

This function parses the API's [Web Services Description Language (WSDL)](https://en.wikipedia.org/wiki/Web_Services_Description_Language) file to determine
possible URL parameters.

Use the following code to get a tibble per API endpoint containing param details:

```r
c("snapshots",
  "congruences",
  "mutations",
  "classifications") %>%
  {
    rlang::set_names(x = purrr::map(.x = .,
                                    .f = swissmuni:::api_params),
                     nm = .)
  }
```

```{r}
#' Get allowed API URL parameters
#'
#' @param type The API endpoint.
#'
#' @return A [tibble][tibble::tbl_df].
#' @keywords internal
#'
#' @examples
#' # to get all URL params of all 3 API endpoints:
#' library(magrittr)
#'
#' endpoints <- c("snapshots",
#'                "congruences",
#'                "mutations",
#'                "classifications")
#'
#' \donttest{
#' endpoints %>%
#'   purrr::map(getFromNamespace("api_params",
#'                               ns = "swissmuni")) %>%
#'   rlang::set_names(nm = endpoints)}
api_params <- function(type = c("snapshots",
                                "congruences",
                                "mutations",
                                "classifications")) {
  
  rlang::arg_match(type) %>%
    switch("snapshots" = "GetSnapshot",
           "congruences" = "GetCorrespondances",
           "mutations" = "GetMutations",
           "classifications" = "GetGeographicLevel") %>%
    paste0("//xs:element[@name='", ., "']") %>%
    xml2::xml_find_all(x = xml2::read_xml("https://sms.bfs.admin.ch/WcfBFSSpecificService.svc?singleWsdl")) %>%
    checkmate::assert_list(len = 1L,
                           types = "xml_node",
                           any.missing = FALSE) %>%
    xml2::xml_find_all(xpath = ".//xs:element") %>%
    xml2::xml_attrs() %>%
    purrr::map(as.list) %>%
    purrr::map_dfr(tibble::as_tibble)
}
```

## Parse API result

**Notes:**

- The API response doesn't have the proper content type set (would be `text/csv`).

- The API returns different numbers of columns depending on the provided URL params (like the `start/end_date`). Therefore we generate the column type spec
  telling `readr::read_csv()` how to parse the results dynamically.

```{r}
parse_result <- function(response,
                         col_types) {
  
  response %<>% httr::content(as = "text")
  
  response %>% readr::read_csv(col_types = pal::cols_regex("(Name|_Title|_Text|^ABBREV)" = readr::col_character(),
                                                           "^(MutationDate|ValidFrom|ValidTo)$" = readr::col_date(format = "%d.%m.%Y"),
                                                           .col_names = pal::dsv_colnames(response),
                                                           .default = readr::col_integer()))
}
```

# Get municipality snapshots

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/snapshots?useBfsCode=true&startPeriod=01-01-2018&endPeriod=01-01-2018>

Possible URL parameters:

- `format`
- `escapeChars`
- `startPeriod`
- `endPeriod`
- `useBfsCode`

```{r}
#' Get municipality snapshots
#'
#' This function returns a [tibble][tibble::tbl_df] with snapshots of Swiss municipality data. A snapshot is the state of the municipalities at a specified
#' point in time. Therefore you probably want to leave the default `end_date = start_date`.
#'
#' @includeRmd snippets/fso_vs_historicized_code.Rmd
#'
#' @param start_date The begin of the period the data covers. A [date](base::Date) or a character scalar in the format `"YYYY-MM-DD"`.
#' @param end_date The end of the period the data covers. A [date](base::Date) or a character scalar in the format `"YYYY-MM-DD"`.
#' @param historicized_code By default, the [_FSO commune number_](https://de.wikipedia.org/wiki/Gemeindenummer) is returned. Set to `TRUE` in order to get the
#'   _historicization number_ instead.
#' @param use_cache `r pkgsnip::param_label("use_cache")`
#' @param cache_lifespan `r pkgsnip::param_label("cache_lifespan")` Defaults to 1 day (24 hours).
#'
#' @return `r pkgsnip::return_label("data")`
#' @export
#'
#' @examples
#' snapshots(historicized_code = TRUE)
snapshots <- function(start_date = lubridate::today(),
                      end_date = start_date,
                      historicized_code = FALSE,
                      use_cache = TRUE,
                      cache_lifespan = "1 day") {
  
  if (checkmate::assert_flag(use_cache)) {
    
    pin_name <- pkgpins::call_to_name()
    result <- pkgpins::get_obj(id = pin_name,
                               max_age = cache_lifespan,
                               pkg = pkg)
    fetch <- is.null(result)
    
  } else {
    fetch <- TRUE
  }
  
  if (fetch) {
    
    result <-
      httr::RETRY(verb = "GET",
                  url = api_base_url,
                  path = paste0(api_common_path, "snapshots"),
                  query = list(startPeriod = as_api_date(start_date),
                               endPeriod = as_api_date(end_date),
                               useBfsCode = tolower(checkmate::assert_flag(historicized_code))),
                  times = 5L) %>%
      parse_result()
  }
  
  if (use_cache & fetch) {
    
    pkgpins::cache_obj(x = result,
                       id = pin_name,
                       pkg = pkg)
  }
  
  result
}
```

# Get municipality congruences

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/correspondances?includeUnmodified=true&startPeriod=01-01-2017&endPeriod=01-01-2018>

Possible URL parameters:

- `format`
- `escapeChars`
- `startPeriod`
- `endPeriod`
- `includeUnmodified`
- `includeTerritoryExchange`

```{r}
#' Get municipality congruences
#'
#' This function returns a [tibble][tibble::tbl_df] with congruence data of Swiss municipalities.
#'
#' @inheritParams snapshots
#' @param incl_unmodified Include municipalities where no changes occurred in the defined period.
#' @param incl_territory_exchange Include information (additional rows) about territory exchanges within and between municipalities.
#'
#' @inherit snapshots return
#' @export
#'
#' @examples
#' congruences(start_date = lubridate::today(),
#'             end_date = lubridate::today(),
#'             cache_lifespan = "6 hours")
congruences <- function(start_date = NULL,
                        end_date = NULL,
                        incl_unmodified = TRUE,
                        incl_territory_exchange = FALSE,
                        use_cache = TRUE,
                        cache_lifespan = "1 day") {
  
  if (checkmate::assert_flag(use_cache)) {
    
    pin_name <- pkgpins::call_to_name()
    result <- pkgpins::get_obj(id = pin_name,
                               max_age = cache_lifespan,
                               pkg = pkg)
    fetch <- is.null(result)
    
  } else {
    fetch <- TRUE
  }
  
  if (fetch) {
    
    result <-
      httr::RETRY(verb = "GET",
                  url = api_base_url,
                  path = paste0(api_common_path, "correspondances"),
                  query = list(startPeriod = as_api_date(start_date),
                             endPeriod = as_api_date(end_date),
                             includeUnmodified = tolower(checkmate::assert_flag(incl_unmodified)),
                             includeTerritoryExchange = tolower(checkmate::assert_flag(incl_territory_exchange))),
                  times = 5L) %>%
      parse_result()
  }
  
  if (use_cache & fetch) {
    
    pkgpins::cache_obj(x = result,
                       id = pin_name,
                       pkg = pkg)
  }
  
  result
}
```

# Get municipality mutations

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/mutations?startPeriod=01-01-2017&endPeriod=01-01-2018&includeTerritoryExchange=false>

Possible URL parameters:

- `format`
- `escapeChars`
- `startPeriod`
- `endPeriod`
- `includeTerritoryExchange`

```{r}
#' Get municipality mutations
#'
#' This function returns a [tibble][tibble::tbl_df] with mutation data of Swiss municipalities.
#'
#' @inheritParams snapshots
#' @inheritParams congruences
#'
#' @inherit snapshots return
#' @export
#'
#' @examples
#' mutations(start_date = "2020-01-01",
#'           end_date = "2020-06-30",
#'           incl_territory_exchange = TRUE)
mutations <- function(start_date = NULL,
                      end_date = NULL,
                      incl_territory_exchange = FALSE,
                      use_cache = TRUE,
                      cache_lifespan = "1 day") {
  
  if (checkmate::assert_flag(use_cache)) {
    
    pin_name <- pkgpins::call_to_name()
    result <- pkgpins::get_obj(id = pin_name,
                               max_age = cache_lifespan,
                               pkg = pkg)
    fetch <- is.null(result)
    
  } else {
    fetch <- TRUE
  }
  
  if (fetch) {
    
    result <-
      httr::RETRY(verb = "GET",
                  url = api_base_url,
                  path = paste0(api_common_path, "mutations"),
                  query = list(startPeriod = as_api_date(start_date),
                               endPeriod = as_api_date(end_date),
                               includeTerritoryExchange = tolower(checkmate::assert_flag(incl_territory_exchange))),
                  times = 5L) %>%
      parse_result()
  }
  
  if (use_cache & fetch) {
    
    pkgpins::cache_obj(x = result,
                       id = pin_name,
                       pkg = pkg)
  }
  
  result
}
```

# Get geographical classifications of municipalities

Example URL:

<https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/levels?startPeriod=01-01-2017&endPeriod=01-01-2018>

Possible URL parameters:

- `format`
- `escapeChars`
- `startPeriod`
- `endPeriod`
- `useBfsCode`
- `labelLanguages`
- `oracleCompatible`

```{r}
#' Get geographical classifications of municipalities
#'
#' This function returns a [tibble][tibble::tbl_df] with a wide range of geographical classifications of Swiss municipalities.
#'
#' There are some classifications which have multiple column IDs (e.g. `HR_ARBREG2000_L2` and `HR_TYPMSR2000_L2` both contain the _spatial mobility regions_).
#' If `name_type` is set to a language code, column duplicates are given the name suffix "`r paste0(' (', cp_symbol, ')')`".
#'
#' @includeRmd snippets/fso_vs_historicized_code.Rmd
#'
#' @inheritParams snapshots
#' @param name_type The column name type to be returned. Defaults to the column IDs returned by the web service. If set to a language code, the columns
#'   starting with ID `HR_` will be named by its full label in the respective language instead.
#'
#' @inherit snapshots return
#' @export
#'
#' @examples
#' classifications(start_date = lubridate::today(),
#'                 end_date = lubridate::today(),
#'                 name_type = "de",
#'                 use_cache = FALSE)
classifications <- function(start_date = NULL,
                            end_date = NULL,
                            historicized_code = FALSE,
                            name_type = c("ID", "en", "de", "fr", "it"),
                            use_cache = TRUE,
                            cache_lifespan = "1 day") {
  
  if (checkmate::assert_flag(use_cache)) {
    
    pin_name <- pkgpins::call_to_name(exclude_args = c("name_type",
                                                       "use_cache",
                                                       "cache_lifespan"))
    result <- pkgpins::get_obj(id = pin_name,
                               max_age = cache_lifespan,
                               pkg = pkg)
    fetch <- is.null(result)
    
  } else {
    fetch <- TRUE
  }
  
  if (fetch) {
    
    result <-
      httr::RETRY(verb = "GET",
                  url = api_base_url,
                  path = paste0(api_common_path, "levels"),
                  query = list(startPeriod = as_api_date(start_date),
                               endPeriod = as_api_date(end_date),
                               useBfsCode = tolower(checkmate::assert_flag(historicized_code))),
                  times = 5L) %>%
      parse_result()
  }
  
  if (use_cache & fetch) {
    
    pkgpins::cache_obj(x = result,
                       id = pin_name,
                       pkg = pkg)
  }
  
  name_type <- rlang::arg_match(name_type)
  
  if (name_type != "ID") {
    
    hr_ix <- which(stringr::str_detect(colnames(result), "^HR_"))
    hr_ids <- colnames(result) %>% stringr::str_subset("^HR_")
    hr_names <- col_spec_classifications[[paste0("name_", name_type)]][match(hr_ids, col_spec_classifications$column_id)]
    
    # ensure column names are unique (there are different column IDs for the same values)
    # -> add the unicode symbol for copy/paste, multiple times if necessary
    while (anyDuplicated(hr_names)) {
      hr_names[which(duplicated(hr_names))] <- paste0(hr_names[which(duplicated(hr_names))], " ", cp_symbol)
    }
    
    colnames(result)[hr_ix] <- hr_names
  }
  
  result
}
```
