---
editor_options:
  chunk_output_type: console
---

# NOTES

-   This file is *not* included in the [source package](https://r-pkgs.org/package-structure-state.html#source-package) because of the [`.nopurl` suffix in the
    filename](https://rpkg.dev/pkgpurl/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to regenerate the package data.

-   Although the datasets below are saved as ["internal data" in `R/sysdata.rda`](https://r-pkgs.org/data.html#data-sysdata), some of them might still be
    exported and documented (in the main `.Rmd` document), something [not explicitly
    mentioned](https://coolbutuseless.github.io/2018/12/10/r-packages-internal-and-external-data/) in the book [R
    Packages](https://r-pkgs.org/data.html#data-data).

# Define data

## `col_spec_classifications`

Generate API response's full column names

NOTES:

-   Unfortunately, this is only possible for the `HR_*` columns of the `levels` API endpoint.

-   The returned names seem to be right (there are columns with completely different IDs like `HR_ARBREG2000_L2` and `HR_TYPMSR2000_L2` but identical values),
    except some year numbers. They are corrected by the function below.

-   The `MS` in the full column names returned by `gen_col_spec_classifications()` stands for *mobilité spatiale*, cf.
    <https://www.bfs.admin.ch/bfs/de/home/statistiken/raum-umwelt/nomenklaturen/msreg.assetdetail.415729.html>.

```{r}
gen_col_spec_classifications <- function() {
  
  pal::assert_pkg("readxl")
  
  # download raw XLS file
  tmp_file <- tempfile(pattern = "col_spec_classifications")
  on.exit(unlink(tmp_file))
  
  httr::RETRY(verb = "GET",
              url = "https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/levels?format=x",
              httr::write_disk(path = tmp_file,
                               overwrite = TRUE),
              times = 5L)
  
  # tidy up data
  readxl::read_excel(path = tmp_file) %>%
    dplyr::rename_with(.fn =
                         ~ tolower(.x) %>%
                         stringr::str_remove_all(pattern = "[\\[\\]]") %>%
                         stringr::str_replace_all(pattern = "\\s+",
                                                  replacement = "_")) %>%
    ## complete duplicated names with missing year
    dplyr::rowwise() %>%
    dplyr::mutate(dplyr::across(dplyr::starts_with("name_"),
                                ~ if (stringr::str_detect(column_id, "^HR_(AGGLGK20(00|12)_L1|DEGURB|MSREG|SPRGEB|STALAN)")) {
                                  paste(.x, stringr::str_extract(column_id, "\\d{4}"))
                                } else .x)) %>%
    ## fix remaining duplicated names
    dplyr::mutate(dplyr::across(dplyr::starts_with("name_"),
                                ~ if (stringr::str_detect(column_id, "^HR_GDETYP(1980|1990|2000)_L[12]$")) {
                                  stringr::str_replace(.x, "1980-2000", stringr::str_extract(column_id, "\\d{4}"))
                                } else .x)) %>%
    dplyr::ungroup()
}

col_spec_classifications <- gen_col_spec_classifications()
```

## `pkg_config`

```{r}
pkg_config <-
  tibble::tibble(key = character(),
                 default_value = list(),
                 description = character()) %>%
  tibble::add_row(key = "global_max_cache_age",
                  default_value = list("30 days"),
                  description = pkgsnip::md_snip("opt_global_max_cache_age"))
```

# Write data

Save all the small data objects as a single internal file `R/sysdata.rda`. Note that when documenting them, they must be explicitly `@export`ed to be available
to package users.

```{r}
usethis::use_data(col_spec_classifications,
                  pkg_config,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```
