---
output: pal::gitlab_document
---

# `r paste0(pal::desc_value("Package"), ": ", pal::desc_value("Title"))`

swissmuni provides functions to access municipality [snapshots](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/snapshots),
[congruences](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/correspondances),
[mutations](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/mutations) and their
[geographical classifications](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/levels) from the
[web services](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc?singleWsdl) provided by the
[Swiss Federal Statistical Office (FSO)](https://www.bfs.admin.ch/bfs/en/home.html). The accessed web services are part of the FSO's
_Statistical Metadata System (SMS)_[^sms].

Note that there's also an official web application [available](https://www.agvchapp.bfs.admin.ch/de/communes/query) to access Swiss municipality data.


[^sms]: Publicly accessible information about this system is scarce. A presentation introducing the system at the
[4th SDMX Global Conference 2013](https://sdmx.org/?sdmx_events=4th-sdmx-global-conference) is found
[here](https://web.archive.org/web/20200615113441/https://www.oecd.org/sdd/SDMX%202013%20Session%203.7%20-%20A%20statistical%20metadata%20system%20based%20on%20SDMX.pdf).


## Installation
```{r, install-snippet-dev, child = pkgsnippets::snippet_path("installation-notice_dev-version_gitlab.Rmd")}
```

## Development

### Literate Programming
```{r, literate-programming-snippet, child = pkgsnippets::snippet_path("literate-programming-notice.Rmd")}
```

### Coding Style
```{r, coding-style-snippet, child = pkgsnippets::snippet_path("coding-style-notice.Rmd")}
```

## See also

- [Official municipality data web application from the Swiss Federal Statistical Office (FSO)](https://www.agvchapp.bfs.admin.ch/de/communes/query)

---
