# swissmuni: Download (meta)data about Swiss municipalities from the [web services](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc?singleWsdl) provided by the [Federal Statistical Office](https://www.bfs.admin.ch/bfs/en/home.html)
```{r, prep-snippets, echo = FALSE}
# download the snippets to insert later on
## define temporary file paths
file_path_installation_notice_dev_version <- tempfile(pattern = "installation_notice_dev_version_")
file_path_literate_programming_notice <- tempfile(pattern = "literate_programming_notice_")

## download the files
download.file(url = "https://gitlab.com/salim_b/snippets/raw/master/R%20Markdown/installation_notice_dev_version.Rmd",
              destfile = file_path_installation_notice_dev_version,
              quiet = TRUE)
download.file(url = "https://gitlab.com/salim_b/snippets/raw/master/R%20Markdown/literate_programming_notice.Rmd",
              destfile = file_path_literate_programming_notice,
              quiet = TRUE)
```
swissmuni is an R package that provides functions to access municipality [snapshots](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/snapshots), [correspondances](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/correspondances) and [mutations](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/mutations) from the [web services](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc?singleWsdl) provided by the [Federal Statistical Office (FSO)](https://www.bfs.admin.ch/bfs/en/home.html). The accessed web services are part of the FSO's _Statistical Metadata System (SMS)_.

## Features

...

## Details

...

## Installation
```{r, install-snippet, child = file_path_installation_notice_dev_version}
```

## Development
```{r, development-snippet, child = file_path_literate_programming_notice}
```
