---
output: pal::gitlab_document
---

# `r paste0(pal::desc_value("Package"), ": ", pal::desc_value("Title"))`

`r paste0('[![CRAN Status](https://r-pkg.org/badges/version/', pal::desc_value(key = "Package"), ')](https://cran.r-project.org/package=', pal::desc_value(key = "Package"), '){.pkgdown-release}')`

swissmuni provides access to Swiss municipality [snapshots](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/snapshots),
[congruences](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/correspondances),
[mutations](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/mutations) and their [spatial
classifications](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/levels) from the [web
services](https://www.bfs.admin.ch/bfs/de/home/dienstleistungen/forschung/api/api-gemeinde.assetdetail.15224054.html) provided by the [Swiss Federal Statistical
Office (FSO)](https://www.bfs.admin.ch/bfs/en/home.html). The accessed web services are part of the FSO's *Statistical Metadata System (SMS)*[^1].

Note that there's also an official web application [available](https://www.agvchapp.bfs.admin.ch/de/communes/query) to access Swiss municipality data.

```{r}
#| label: documentation
#| eval: !expr '!isTRUE(getOption("pal.build_readme.is_pkgdown"))'
#| results: asis
#| echo: false
pkgsnip::md_snip(name = "pkgdown_notice") %>%
  paste0("## Documentation\n\n",
         "[![Netlify Status](https://api.netlify.com/api/v1/badges/9c255431-9f12-4bd9-a169-a7be1e23b985/deploy-status)]",
         "(https://app.netlify.com/sites/swissmuni-rpkg-dev/deploys)\n\n",
         .) %>%
  pal::cat_lines()
```

[^1]: Publicly accessible information about this system is scarce. A presentation introducing the system at the [4th SDMX Global Conference
    2013](https://sdmx.org/?sdmx_events=4th-sdmx-global-conference) is found
    [here](https://web.archive.org/web/20200615113441/https://www.oecd.org/sdd/SDMX%202013%20Session%203.7%20-%20A%20statistical%20metadata%20system%20based%20on%20SDMX.pdf).

## Installation

```{r}
#| label: install-snippet-dev
#| child: !expr pkgsnip::snip_path("installation-notice_dev-version_gitlab.Rmd")
```

```{r}
#| label: usage
#| eval: !expr isTRUE(getOption("pal.build_readme.is_pkgdown"))
#| results: asis
#| echo: false
pkgsnip::md_snip(name = "usage_notice") %>%
  paste0("## Usage\n\n", .) %>%
  pal::cat_lines()
```

## Development

### R Markdown format

```{r}
#| label: rmd-format
#| child: !expr pkgsnip::snip_path("rmd-package-notice.Rmd")
```

### Coding style

```{r}
#| label: coding-style
#| child: !expr pkgsnip::snip_path("coding-style-notice.Rmd")
```

## See also

-   [Official municipality data web application from the Swiss Federal Statistical Office (FSO)](https://www.agvchapp.bfs.admin.ch/de/communes/query)
