---
output: pal::gitlab_document
---

# `r paste0(pal::desc_value("Package"), ": ", pal::desc_value("Title"))`

swissmuni provides functions to access municipality [snapshots](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/snapshots), [correspondances](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/correspondances) and [mutations](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/mutations) from the [web services](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc?singleWsdl) provided by the [Swiss Federal Statistical Office (FSO)](https://www.bfs.admin.ch/bfs/en/home.html). The accessed web services are part of the FSO's _Statistical Metadata System (SMS)_.

## Installation
```{r, install-snippet-dev, child = pkgsnippets::snippet_path("installation-notice_dev-version_gitlab.Rmd")}
```

## Development

### Literate Programming
```{r, literate-programming-snippet, child = pkgsnippets::snippet_path("literate-programming-notice.Rmd")}
```

### Coding Style
```{r, coding-style-snippet, child = pkgsnippets::snippet_path("coding-style-notice.Rmd")}
```

## See also

- ...
