---
output: pal::gitlab_document
---

# `r paste0(pal::desc_value("Package"), ": ", pal::desc_value("Title"))`

swissmuni provides access to Swiss municipality [snapshots](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/snapshots),
[congruences](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/correspondances),
[mutations](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/mutations) and their [spatial
classifications](https://sms.bfs.admin.ch/WcfBFSSpecificService.svc/AnonymousRest/communes/levels) from the [web
services](https://www.bfs.admin.ch/bfs/de/home/dienstleistungen/forschung/api/api-gemeinde.assetdetail.15224054.html) provided by the [Swiss Federal Statistical
Office (FSO)](https://www.bfs.admin.ch/bfs/en/home.html). The accessed web services are part of the FSO's *Statistical Metadata System (SMS)*[^1].

Note that there's also an official web application [available](https://www.agvchapp.bfs.admin.ch/de/communes/query) to access Swiss municipality data.

```{r, pkgdown, eval = !isTRUE(getOption("pal.build_readme.is_pkgdown")), results = "asis", echo = FALSE}
pkgsnip::md_snip(name = "pkgdown_notice") %>%
  paste0("## Documentation\n\n", .) %>%
  pal::cat_lines()
```

[^1]: Publicly accessible information about this system is scarce. A presentation introducing the system at the [4th SDMX Global Conference
    2013](https://sdmx.org/?sdmx_events=4th-sdmx-global-conference) is found
    [here](https://web.archive.org/web/20200615113441/https://www.oecd.org/sdd/SDMX%202013%20Session%203.7%20-%20A%20statistical%20metadata%20system%20based%20on%20SDMX.pdf).

## Installation

```{r, install-snippet-dev, child = pkgsnip::snip_path("installation-notice_dev-version_gitlab.Rmd")}
```

## Development

### R Markdown format

```{r, rmd-pkg-snippet, child = pkgsnip::snip_path("rmd-package-notice.Rmd")}
```

### Coding style

```{r, coding-style-snippet, child = pkgsnip::snip_path("coding-style-notice.Rmd")}
```

## See also

-   [Official municipality data web application from the Swiss Federal Statistical Office (FSO)](https://www.agvchapp.bfs.admin.ch/de/communes/query)
